### 2014-09-15 PrivacyProtection-2014-09-09.tex
WIP reviews model

The \emph{reference model} consists of a \emph{model}, which aims to organize common security terminology in a reasonable and practical way and a systematic analysis \emph{procedure} based on that model.

In this section we give a brief overview over the reference model.
Before we can start with the description of model and procedure we introduce the following basic terminology (cf. Figure \ref{fig:terminology}).

%\includegraphics{diagrams/png/itsecterminology.png}
\input{./figures/Basic-IT-Security-Terminology}

\item A \emph{threat} is a class of potential events whose manifestation can
cause damage or harm.

\item A \emph{vulnerability} is a weakness which can cause the loss of security.

\item a \emph{hazard} is a concrete risk due to one or more concrete vulnerabilities and corresponding threats.

\texttt{threat + vulnerability -\textgreater{} hazard}

\item An \emph{incident} is an event caused by a threat manifestation against one or more
corresponding vulnerabilities.

\texttt{threat manifestation + vulnerability -\textgreater{} hazard}

\item An \emph{accident} is an incident which did cause damage or harm. (Translates to 'Schandensfall/Zwischenfall/Störfall' in the German language).

\item A \emph {non-accident} is an incident which did not cause any damage or harm.

\item A security \emph{measure} is a set of one or more activities which reduce or eliminate
vulnerabilities.

@@

The \textbf{world view} contains all components describing the current state.
It consists of the following components:
\begin{itemize}
\item
\textbf{Assets.}
Things of value to one or more stakeholders.
The value can be ``hard'' (money, data, etc.) or ``soft'' (trust,privacy,etc.).

\item \textbf{IT-Systems.}
The IT-Systems under study.

\item \textbf{Humans.}
All identifiable actors of the system under study.
Many problems arise due to misunderstandings during man-machine interaction.

\item \textbf{Conflicts of Interests.}
Different actors have different interests.
Those interest can be in conflict.
A trival conflict is the \emph{Attacker-Attackee-Conflict}:
A service provider offers private data storage, therefore the provider is interested in having the access restricted.
An attacker is naturally interested in easy access in order captialize the stolen data.
However, heavy security restriction are also a burden for the service operator, as they are interested in efficient work flows.
This kind of conflict can create vulnerabilities.

\item \textbf{Vulnerabilities.}
All indentifiable weaknesses in the current system layout.

\item \textbf{Interactions.}
Assets, IT-Systems, Humans and Vulnerabilites are in continous interaction with each other in order to support the business objectives.
Those interactionscan also lead to threats, e.g. having unencrypted communication with a server.
Therefore all interactions which occur in the system's outline have to be documented.
\end{itemize}

The \textbf{potential view} describes both the desired and dreaded state.
It consists of the following components.
\begin{itemize}
\item \textbf{Chances/Risk.}
Chances and Risks affect business objectives and threats.

\item \textbf{Business Objectives.}
Chances the system under study.
This describes the desired state.

\item \textbf{Threats.}
Risks for business objectives, this can simply be the integrety of the IT-System.
Threats target one or more assets and violate security requirements}.
This describes the dreaded state.

\item
\textbf{Security Requirements:} A set of distinct requirements for a safe and secure system.
Security requirements hold for one or more assets.
This describes the potential state after this procedure.
\end{itemize}

The \textbf{concept view} is a realization of the the potential view of the system.
It describes a plan of sufficient measures to secure the desired state and to aviod the dreaded state of the system.
It contians the following components:
\begin{itemize}
\item \textbf{Business Model.}
The plan to achieve business objectives.

@@

\item \textbf{Attack Technique.}
A specific technique or technology to attack IT-Systems (Man in the Middel, Phishing, etc.).
(\textbf{Note:} This component is not conform to the glossary as the term \emph{accident} there covers both disaster and attack!)

\item \textbf{Security Measures.}
The plan to achieve a safe and secure system.
Each security mesaure opposes an accident scenario or an attack technique.

The \textbf{event view} contains all actual operations and events through out the production phase.
The event view instantiates the concept view of the system.
It contains the following components:
\begin{itemize}
\item
  \textbf{Business Process:} The actual, running instance of the
  \textbf{Busieness Modell}
\item
  \textbf{Accidents:} All actually happened accidents.
\item
  \textbf{Attacks:} All actually happened attacks. (\textbf{Note:} This
  component is not conform to the glossary as the term \emph{accident}
  there covers both disaster and attack!)
\item
  \textbf{Security Operations:} Instances of \textbf{Security Measures}.
  Security Measures \emph{secure} the system against \textbf{Accidents}
  and \textbf{Attacks}.
\end{itemize}

>>>>

The reference model consists of a \emph{model} and a \emph{procedure}.
The model aims to organize common security terminology in a reasonable and practical way.
The procedure describes a systematic way to analysis the IT system based on that model.
In this section we give a brief overview over the reference model.

@@

The \textbf{world view} contains all components describing the current state.
It consists of the following components:
\begin{itemize}
\item
\textbf{Assets.} Things of value to one or more stakeholders.
The value can be ``hard'' (money, data, etc.) or ``soft'' (trust, privacy,etc.).
In our case the only asset we are concerned with is the privacy of the citizen.

\item \textbf{IT-Systems.}
A description of the IT-Systems under study.
The level of garnularity has to be detailed enough to express all possible threats to the assets at stake.

\item \textbf{Actors.}
All identifiable stakeholders of the system under study.
Typical actors include, users, developers, clients and criminals.

\item \textbf{Conflicts of Interests.}
Different actors have different interests which can be in conflict which each other.
These conflicts of interest are the origin of all attacks to the system.

A typical conflict is the \emph{Criminal-User-Conflict}:
A user wants to keep control over their private data.
A criminal wants to gain money.
The possibility of selling private data (user profiles) to advertizers, renders both interests conflicting.

\item \textbf{Vulnerabilities.}
All indentifiable weaknesses in the IT-System.

In the example of the cirminal-user-conflict, the criminal has to exploit a vulnerability, e.g. a weak password, to gain access to the private data abou the user.

\item \textbf{Interactions.}
This point captures all possible interactions between assets, IT-Systems, humans and vulnerabilites.
It is described in more detail in the next view.
\end{itemize}

The \textbf{potential view} describes the intended and unintended interactions of the components of the world view.
The intended interactions support the underlying business objectives.
Unintended interactions lead to threats.
The potential view consists of the following components.

\begin{itemize}
\item \textbf{Business Objectives.}
Describes how IT-system and actors should interact in order to reach the goals of operation.

\item \textbf{Threats.}
A threat is a potential interaction that destroys or harms assets of the system.
Concrete realizations of threats can be \emph{attacks} or \emph{accidents}.
Attacks are executed by an actor in response to a conflict of interest.
Accidents are harmful interactions that are not willfully caused by an actor.

\item \textbf{Chances/Risk.}
Evaluation of chances and risks associated to the business objectives and threats.
The risk associated to a threat is its expected loss.
A chance associated to a intended interaction is its expected gain.

In the case that, the loss can be quantified monetary, and the likelihood of occurence of a threat can
be modeled probabilistically, the risk is given by the procuct
\[ \text{risk} = \text{loss} \cdot P[\text{threat}]. \]
In practice such a quantitative risk evaluation is often not possible, and a qualitative, heursitic, analysis is performed instead.

\item \textbf{Security Requirements.}
Targeted change in specify interactions that may not occure within the system in order to acieve its business objectives.
Security requirements are targeted to protect one or more assets.

The \textbf{concept view} is a realization of the the potential view of the system.
It describes a plan of sufficient measures to secure the intended interactions and to aviod the unintended interactions.
It contians the following components:
\begin{itemize}
\item \textbf{Business Model.}
The plan to achieve business objectives.

@@

\item \textbf{Attack Technique.}
A specific technique or technology to attack IT-Systems (Man in the Middel, Phishing, etc.).

%TODO: (\textbf{Note:} This component is not conform to the glossary as the term \emph{accident} there covers both disaster and attack!)

\item \textbf{Security Measures.}
The plan to achieve a safe and secure system.
Each security mesaure opposes an accident scenario or an attack technique.
\end{itemize}

The \textbf{event view} contains all actual events through out the lifetime of the system.
The event view instantiates the concept view of the system.
It contains the following components:
\begin{itemize}
\item
  \textbf{Business Process:} The actual, running instance of the business model.
\item
\textbf{Accidents:} All actually happened accidents.

\item
\textbf{Attacks:} All actually happened attacks. (\textbf{Note:} This
component is not conform to the glossary as the term \emph{accident}
there covers both disaster and attack!)

\item
\textbf{Security Operations:} Instances of \textbf{Security Measures}.
Security Measures \emph{secure} the system against \textbf{Accidents}
and \textbf{Attacks}.
\end{itemize}
